---
title: "Age of cited authorities"
format: 
  html:
    code-fold: true
    df-print: kable
    embed-resources: true
---

We have a list of treatises (well, so far, volumes in treatises) that cite specific cases. We want to understand how old the cases were when a treatise cited them.

```{r}
#| output: false
library(tidyverse)
library(DBI)
library(dbplyr)
db <- dbConnect(odbc::odbc(), "Research", timeout = 10)
```

Here are all citations in U.S. treatises to U.S. case law that we can link. This shows a sample of what the data looks like.

```{r}
citations <- tbl(db, in_schema("linking", "moml_06_treatise_to_case"))
treatises <- tbl(db, in_schema("moml", "us_treatises")) |> mutate(psmid = as.character(psmid))
cases <- tbl(db, in_schema("cap", "cases"))
subjects <- tbl(db, in_schema("moml", "book_subject")) |>
  select(-source) |>
  filter(subject != "US", subject != "UK") |>
  collect()

citations <- citations |>
  left_join(treatises |> select(psmid, treatise_year = year), by = c("moml_treatise" = "psmid")) |>
  left_join(cases |> select(id, case_year = decision_year), by = c("case" = "id")) |>
  select(-n) |>
  rename(treatise = moml_treatise) |>
  filter(treatise_year >= case_year) |>
  collect()
head(citations, 10)
```

We should group the treatises by decade. And we can compute the age of the citations.

```{r}
decade <- function(x) {
  trunc(x / 10, 0) * 10
}

citations <- citations |> 
  mutate(treatise_decade = decade((treatise_year))) |>
  mutate(cite_age = treatise_year - case_year)

head(citations, 10)
```

Here is a table of the average age of the case cited in treatises published in a certain year. 

```{r}
avg_age <- citations |>
  group_by(treatise_year) %>%
  summarize(
    avg_age = cite_age |> mean() |> round(1),
    n = n()) 
avg_age
```

Since some of those data points are outliers because there are so few treatises, we will exclude them when making a plot.

```{r}
#| warning: false
avg_age |>
  filter(n > 100) |>
  ggplot(aes(x = treatise_year, y = avg_age)) + 
  geom_point() +
  geom_line() +
  geom_smooth(se = FALSE, color = "red") + 
  labs(
    x = "Year treatise published",
    y = "Average age of cases cited",
    title = "Age of cases cited in U.S. treatises"
  ) +
  theme_light()
```

We can instead visualize this as a boxplot per decade.

```{r}
citations |>
ggplot(aes(x = factor(treatise_decade), y = cite_age)) +
geom_boxplot(outlier.shape = NA) +
  labs(
    x = "Decade treatise published",
    y = "Age of cases cited",
    title = "Age of cases cited in U.S. treatises"
  ) +
  ylim(0, 100) +
  theme_light()
```

Let's try to split this by topic. We only want to keep some of them.

```{r}
subjects |>
count(subject, sort = TRUE) %>% 
head(20)
```

Now we can plot the change over time by subject.

```{r}
#| warning: false
subject_keepers <- c("Procedural Law", "Real Property", "Criminal Law",
"Constitutional Law", "Civil Procedure", "Equity", "Trusts and Estates", "Family Law")
citations |>
left_join(subjects |> filter(subject %in% subject_keepers), by = c("treatise" = "psmid"), multiple = 'all') |>
group_by(subject, treatise_year) %>%
summarize(
    avg_age = cite_age |> mean() |> round(1),
    n = n()
) |>
ggplot(aes(x = treatise_year, y = avg_age)) + facet_wrap(~subject, nrow = 3, scales = "free_y") + 
geom_smooth(se = FALSE) + 
geom_point(alpha = 0.25) +
labs(title = "Average age of cited cases by subject of treatise",
x = "Year treatise published", y = "Average age of cases cited")
```


```{r}
#| output: false
dbDisconnect(db)
```
